% ARR Responsible NLP Research Checklist
% Reference: https://aclrollingreview.org/responsibleNLPresearch/

\section*{Responsible NLP Research Checklist}

This paper follows the ARR Responsible NLP Research checklist.

\subsection*{A. For Every Submission}

\paragraph{A1. Did you describe the limitations of your work?}
Yes. Section~\ref{sec:conclusion} (Limitations paragraph) discusses: limited seed coverage for repo-level comparisons, synthetic-to-real transfer gap, lack of downstream re-repair validation, and val-test mutation distribution mismatch. Section~\ref{sec:repo} (Limitations paragraph) discusses val-best checkpoint selection variance and seed coverage.

\paragraph{A2. Did you discuss any potential risks of your work?}
This work compares formulations for execution-free code verification on public benchmarks. The primary application is improving automated program repair pipelines. We identify no significant risks: the methods predict test outcomes rather than generating executable code, the datasets are derived from open-source repositories, and no personally identifiable information is involved. Misuse potential is limited because the verifiers require task-specific fine-tuning on labeled data and do not generalize to arbitrary code generation.

\subsection*{B. Did You Use or Create Scientific Artifacts?}

\paragraph{B1. Did you cite the creators of artifacts you used?}
Yes. All datasets (HumanEval~\cite{chen2021humaneval}, MBPP~\cite{austin2021mbpp}, SWE-bench Verified~\cite{chowdhury2024swebenchverified}), models (Qwen3~\cite{qwen2025qwen3}, DeepSeek-Coder~\cite{guo2024deepseekcoder}), and methods (LoRA~\cite{hu2022lora}, QLoRA~\cite{dettmers2024qlora}) are cited in Sections~\ref{sec:setup} and~\ref{sec:formulations}.

\paragraph{B2. Did you discuss the license or terms for use and/or distribution of any artifacts?}
HumanEval is released under MIT license; MBPP under CC-BY-4.0; SWE-bench Verified under MIT license. Qwen3 models are released under Apache 2.0; DeepSeek-Coder under the DeepSeek License. All uses are consistent with their respective licenses. We do not release new artifacts in this submission.

\paragraph{B3. Did you discuss if your use of existing artifact(s) was consistent with their intended use?}
Yes. All benchmarks are used for their intended purpose of evaluating code understanding and generation capabilities. SWE-bench Verified is used for repository-level evaluation as designed by its creators.

\paragraph{B4. Did you discuss the steps taken to check whether the data contains any information that names or uniquely identifies individual people or offensive content?}
Not applicable. Our datasets consist of code snippets and test cases from open-source repositories. The mutation-based data construction process operates on source code only and does not involve personal data or natural language content that could contain offensive material.

\paragraph{B5. Did you provide documentation of the artifacts?}
Yes. Section~\ref{sec:setup} documents dataset statistics (sizes, splits, label distributions, majority baselines). The mutation-based data construction process is described in Section~\ref{sec:setup}. Appendix~\ref{app:hyperparams} provides full hyperparameter details.

\paragraph{B6. Did you report relevant statistics like the number of examples, details of train/test/dev splits, etc.?}
Yes. Section~\ref{sec:setup}: function-level 19,076 samples (train 15,290 / val 1,896 / test 1,890; pass:fail $\approx$ 52:48); repo-level ${\sim}$70K training, val $n{=}9{,}335$, test $n{=}9{,}784$ (majority baseline 72.96\%).

\subsection*{C. Did You Run Computational Experiments?}

\paragraph{C1. Did you report the number of parameters in the models used, the total computational budget, and computing infrastructure?}
Yes. Models: Qwen3-4B (${\sim}$4B parameters, LoRA rank 16 $\approx$ 4M trainable), Qwen3-8B (${\sim}$8B), DeepSeek-Coder-6.7B (${\sim}$6.7B). Infrastructure: 4$\times$ NVIDIA RTX 5090 GPUs. Total compute: ${\sim}$370 GPU-hours across 223 completed runs (${\sim}$46 configurations including seeds, formulations, model sizes, and ablations). Appendix~\ref{app:hyperparams} provides per-run details.

\paragraph{C2. Did you discuss the experimental setup, including hyperparameter search and best-found hyperparameter values?}
Yes. Section~\ref{sec:setup} reports the training configuration (LoRA rank 16, 3 epochs, 5 seeds with mixed precision). Appendix~\ref{app:hyperparams} provides full hyperparameter values. We use a fixed configuration rather than extensive hyperparameter search; repo-level models use validation accuracy for checkpoint selection.

\paragraph{C3. Did you report descriptive statistics about your results with, e.g., error bars or confidence intervals?}
Yes. All multi-seed results report mean $\pm$ standard deviation. Confidence intervals use clustered bootstrap (114 problem clusters, 10K resamples; Section~\ref{sec:setup}). Equivalence testing uses TOST ($p = 0.007$, $\pm$2\pp{} margin; Section~\ref{sec:equivalence}). ICC quantifies within-problem correlation with design effect reported.

\paragraph{C4. If you used existing packages, did you report the implementation and version numbers?}
Yes. Implementation uses HuggingFace Transformers (${\geq}$4.44), PEFT (${\geq}$0.12), PyTorch (${\geq}$2.0), bitsandbytes (${\geq}$0.43, for NF4 quantization), and Python 3.10. LoRA~\cite{hu2022lora} and QLoRA~\cite{dettmers2024qlora} configurations are detailed in Appendix~\ref{app:hyperparams}. Exact version pinning is provided in the code repository.

\subsection*{D. Did You Use Human Annotators?}

Not applicable. This work uses no human annotators or participants. All labels are derived from automated test execution.

\subsection*{E. AI Assistants}

\paragraph{E1. Did you use AI assistants (e.g., ChatGPT, Copilot) in any part of the paper writing or coding process?}
Yes. AI coding assistants were used during development of training and evaluation scripts. AI writing assistants were used for drafting and editing portions of the manuscript. All scientific claims, experimental design, and analysis were conducted and verified by the authors.
