\section{Conclusion}
\label{sec:conclusion}

We presented the first controlled comparison of three formulations for execution-free code repair verification: per-test binary classification (\cls{}), execution output generation (\cwm{}), and trajectory-level scoring. Our central finding is that no single formulation dominates across all conditions; the design space is governed by a formulation $\times$ model $\times$ complexity interaction.

At function level, \cls{} and \cwm{} achieve near-equivalent accuracy (gap 1.24\pp{} at 4B, TOST equivalence $p{=}0.007$); formulation choice matters less than model capacity and training data at this complexity tier. At repository level, formulation divergence amplifies dramatically (4B: 25.59\pp{}; 8B: 19.31\pp{}). Cross-model validation with DeepSeek-Coder confirms this divergence is formulation-inherent: both model families show \cwm{} $\gg$ \cls{} degradation from function to repo level. The magnitude, however, is model-contingent (Qwen3: ${\sim}$5/22\pp{}; DeepSeek: ${\sim}$12/25\pp{} for \cls{}/\cwm{}). %%SCENARIO-A%% \cwm{} additionally exhibits higher seed sensitivity (${\sim}$5\pp{} across seeds vs.\ \cls{}'s ${\sim}$2\pp{}), though epoch 2 and epoch 3 achieve comparable accuracy, indicating no overtraining under our protocol. Evaluation protocol itself introduces ${\sim}$9\pp{} accuracy variation for the same \cwm{} checkpoint, a side-effect of the unparseable rate $\to$ majority-class default chain that practitioners must control for. %%END-SCENARIO-A%%

Two formulation-level properties generalize independent of model choice. First, per-test classification provides diagnostic granularity at equivalent trajectory-level accuracy, identifying which specific tests a patch fails. This advantage concentrates in instances with $\geq$2 fail-to-pass tests (31\% of instances, contributing 55\% of total oracle advantage). Second, the 1-token classification formulation yields a 217$\times$ throughput advantage over variable-length generation, an inherent property of the output-length asymmetry.

Error complementarity between formulations (oracle ensemble: 90.63\%) suggests that the practitioner's choice need not be exclusive. Classification and generation exhibit complementary failure patterns. This motivates routing strategies that select formulations based on predicted instance complexity.

\paragraph{Limitations.}
Repo-level cross-model \cwm{} results are preliminary (DeepSeek-Coder \cwm{} training in progress); \cls{} results are final. The repo-level same-size comparisons each use a single seed (4B-4B and 8B-8B); multi-seed DeepSeek-Coder \cls{} evaluation is in progress. For \cwm{}, validation loss diverges to NaN (DeepSeek-Coder), preventing standard early stopping and requiring checkpoint selection by validation accuracy. No downstream re-repair experiment validates whether per-test diagnostic signals translate to actual fix improvements. Pass-to-pass regression rates are negligible (1.04\% mean, 0\% median), meaning the per-test advantage currently concentrates in fail-to-pass discrimination; regression detection remains untested in practice. Training data comprises synthetic mutations generated by operator-level transformations (operator swap, off-by-one, return value, boundary removal). While prior work has established that synthetic mutants correlate with real fault detection~\cite{andrews2005mutation,just2014mutants}, the coupling is imperfect: real bugs often involve multi-statement changes, semantic misunderstandings, or integration errors that no single mutation operator captures. Our repo-level evaluation partially mitigates this concern by using SWE-bench Verified test cases derived from real pull requests, but the underlying code changes remain synthetic. Transfer to naturally occurring developer bugs is untested. The oracle analysis assumes ground-truth test labels; real-world per-test advantage is upper-bounded by classifier accuracy.

\paragraph{Future work.}
Cross-model repo-level validation (\cls{} final, \cwm{} preliminary) confirms the formulation-inherent divergence pattern; multi-seed replication with final \cwm{} training runs would strengthen the statistical claims at this complexity tier. Error complementarity motivates complexity-aware routing that selects formulations per-instance. A pilot study with 8B-scale repair agents found no measurable benefit from per-test diagnostics over binary accept/reject feedback; translating diagnostic granularity into repair improvements requires further investigation. Integration with test-time scaling systems (best-of-$N$ with per-test reranking) is a natural application of the diagnostic granularity finding, though the repair-side value remains to be demonstrated.