\section{Introduction}
\label{sec:introduction}

LLM-based automated program repair systems increasingly rely on test-time compute scaling: generating $K$ candidate patches and selecting the best via a verifier~\cite{snell2024scaling}. Full execution verification is reliable but expensive (\$1--3 per candidate at repository scale\footnote{Based on per-instance costs reported by \citet{yang2024sweagent} and \citet{xia2024agentless}.}), driving demand for execution-free alternatives. Three distinct formulations have emerged for this task: \textbf{generation-based simulation}~\cite{meta2025cwm,maimon2026selfexecution} predicts execution output strings per test; \textbf{trajectory-level scoring}~\cite{shum2025swerm} predicts a binary patch verdict without per-test decomposition; and \textbf{per-test classification}~\cite{ni2023lever,chen2023codet} predicts a pass/fail token per test. Each makes different tradeoffs between diagnostic granularity, output complexity, and inference cost.

Despite active development along each axis, no prior work provides a systematic comparison holding model capacity, training data, and compute constant. Practitioners choosing a formulation for their verification pipeline must rely on results from different models, different datasets, and different evaluation protocols. Whether one formulation is uniformly preferable, or whether the choice depends on properties of the code being verified, remains an open empirical question. This connects to a broader pattern in NLP: task formulation---classification vs.\ generation vs.\ span extraction---systematically affects performance~\cite{raffel2020t5}, but the conditions under which each wins remain task-dependent and empirically unresolved. Recent evidence that generation-based verification outperforms classification in mathematical reasoning~\cite{hosseini2025genrm} raises a further question: does this advantage transfer to code verification, where outputs are longer and format-constrained? A related question is whether formulation choice matters at all relative to more basic decisions: at what complexity level does fine-tuning become necessary, and does formulation selection add value beyond what fine-tuning alone provides?

We present the first systematic empirical comparison across two representative code LLM families (Qwen3, DeepSeek-Coder), two complexity levels (function-level and repository-level), and multiple seeds. Our contribution is primarily methodological: a matched experimental framework that controls for major confounding variables. The analysis also produces actionable outputs: a failure mode taxonomy that characterizes when and why each formulation fails (\S\ref{sec:repo}), error complementarity findings with practical ensemble implications (\S\ref{sec:analysis}), and formulation selection recommendations conditioned on code complexity (\S\ref{sec:conclusion}). The results suggest a three-layer structure. Fine-tuning is necessary (0-shot$\to$fine-tuned gains of 35--36 percentage points (pp) at repo level). After fine-tuning under our ${\leq}$8B LoRA protocol, \emph{formulation choice is moderated by code complexity, operating through output format}: function-level outputs average ${\sim}$5--15 tokens (minimizing generation-specific failures per \S\ref{sec:failure-modes}), while repo-level outputs average ${\sim}$477 tokens (triggering format collapse): at function level, all formulations achieve comparable accuracy---consistent with the theoretical analysis in \S\ref{sec:failure-modes}, which predicts that short outputs minimize generation-specific failures. At repository level under our ${\leq}$8B LoRA protocol, this equivalence diverges; variant-level analysis identifies three distinct failure modes consistent across model families, though gap magnitude depends on multiple confounds (\S\ref{sec:threats}). Figure~\ref{fig:reformulation} illustrates the three formulations.

We make the following contributions:

\begin{itemize}[leftmargin=*,itemsep=2pt]
\item \textbf{Complexity-moderated formulation choice.} We establish a three-layer structure: (1)~fine-tuning is necessary (35--36\pp{} gains); (2)~at function level, formulations are near-equivalent---\cls{} and \cwm{} achieve statistical equivalence (TOST $p{=}0.007$, Holm-corrected) across both families, while \traj{} equivalence is inconclusive after Holm correction, but the gap narrows from 3.11\pp{} at 4B to 1.07\pp{} at 14B (\S\ref{sec:scaling}); (3)~at repository level under ${\leq}$8B LoRA, this equivalence diverges ($z{=}-23.7$, DeepSeek-Coder), with \cwm{} falling 18--26\pp{} below \cls{}, accompanied by format collapse---four interventions all fail to close the gap. The TOST framework demonstrates strict equivalence rather than mere non-significance. Error analysis reveals partially complementary patterns (oracle: 90.63\% function, 95.83\% repo), though simple heuristics capture $<$0.4\pp{}, framing ensemble routing as an open challenge (\S\ref{sec:equivalence}, \S\ref{sec:repo}, \S\ref{sec:analysis}).

\item \textbf{Failure mode taxonomy.} Variant-level analysis identifies three distinct failure modes at repository level---format collapse (\cwm{}), decision collapse (\cls{} under truncate preprocessing), and aggregation penalty (\cls{} under all-pass rule)---providing actionable diagnostic criteria for practitioners selecting a verification formulation (\S\ref{sec:repo}, Table~\ref{tab:taxonomy}).
\end{itemize}
